\documentclass{amsart}
\usepackage[margin=1.1in]{geometry} 
\usepackage{amsmath}
\usepackage{tcolorbox}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{lastpage}
\usepackage{fancyhdr}
\usepackage{accents}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{color}
\input{shortcuts.tex}
\setlength{\headheight}{40pt}


\newenvironment{solution}
  {\renewcommand\qedsymbol{$\blacksquare$}
  \begin{proof}[Solution]}
  {\end{proof}}
\renewcommand\qedsymbol{$\blacksquare$}

\usepackage{amsmath, amssymb, tikz, amsthm, csquotes, multicol, footnote, tablefootnote, biblatex, wrapfig, float, quiver, mathrsfs, cleveref, enumitem, upgreek}
\addbibresource{refs.bib}
\theoremstyle{definition}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{question}[theorem]{Question}
\newtheorem{example}[theorem]{Example}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem*{remark}{Remark}
\newtheorem{definition}[theorem]{Definition}
\numberwithin{equation}{section}
\begin{document}
\large
\title[Algebraic Number Theory]{MATH 223A: Algebraic Number Theory}
\author{Wern Juin Gabriel Ong}
\address{Bowdoin College, Brunswick, Maine 04011}
\email{gong@bowdoin.edu}
\urladdr{https://wgabrielong.github.io/}
\maketitle
\section*{Preliminaries}
These notes roughly correspond to the course \textbf{MATH 223A: Algebraic Number Theory} taught by Prof. Melanie Wood at Harvard University in the Fall 2023 semester. These notes are \LaTeX-ed after the fact with significant alteration and are subject to misinterpretation and mistranscription. Use with caution. Any errors are undoubtedly my own and any virtues that could be ascribed to these notes ought be attributed to the instructor and not the typist. A special thanks goes to Matt King, the official note taker for the course, for his notes have greatly improved my own. 
\tableofcontents
\section{Lecture 1 -- 11th September 2023}
This is a first graduate course in algebraic number theory. We will study local fields. 

Let us understand the what and why of local fields. Local fields are interesting in the context of global fields. Let us consider some examples of global fields. 
\begin{example}[Number Fields]
  $K/\QQ$ where $[K:\QQ]<\infty$, finite extensions of $\QQ$. 
\end{example}
\begin{example}[Function Fields]
  Function fields of curves over finite fields $\FF_{q}$. 
\end{example}
One's first study of number theory seeks to treat $\ZZ\subseteq\QQ$, but there is a strong analogy between studying $\ZZ\subseteq\QQ$ and $\FF_{q}[t]\subseteq\FF_{q}(t)$. In particular, one can show that there is a bijection betwween smooth geometrically irreducible curves over $\FF_{q}$ up to isomorphism and finite extensions of $\FF_{q}(t)$ up to isomorphism. 
\begin{example}
  The field $\FF_{q}(t)$ is the field of rational functions of $\PP^{1}_{\FF_{q}}$. 
\end{example}
This is an example of the function field analogy, drawing connections between geometry over finite fields $\FF_{q}$ and $\CC$. In fact, function fields and number fields are the only examples of global fields. This was first studied in the 20th century through the study of class field theory, field extensions with Abelian Galois group, that saw similar methods applied to number fields and function fields of $C_{/\FF_{q}}$. Artin-Whaples axiomatized what was special about function fields and number fields by defining global fields and valuations. This led to a new approach to proving Dirichlet's unit theorem in the 1940s. Valuations will be our starting point for this class. 
\begin{definition}[Valuation]\label{def:valuation}
  Let $K$ be a field. A valuation on $K$ is a function $|\cdot|:K\to\RR$ such that: 
  \begin{enumerate}[label=(\alph*)]
    \item $|x|\geq0$ for all $x\in K$ and $|x|=0$ if and only if $x=0$, 
    \item $|xy|=|x|\cdot|y|$, 
    \item and $|x+y|\leq|x|+|y|$. 
  \end{enumerate}
\end{definition}
Let us consider some valuations of fields. 
\begin{example}[Trivial Valuation]
  There is a trivial valuation
  $$\begin{cases}
  |x| = 0 & x=0 \\
  |x| = 1 & x\neq0
  \end{cases}$$
  on $K$. 
\end{example}
We can see that both $\RR$ and $\CC$ are fields with the ``standard'' valuations. Let $K$ be a number field. $K$ admits homomorphisms to $\RR$ and/or $\CC$ and inherits the absolute value from $\RR$ and/or $\CC$. More precisely, if $[K:\QQ]=n$, there are $n$ homomorphisms to $\CC$ giving $n$ possibly distinct valuations on $K$. 
\begin{example}
  The field $\QQ(i)$ is of degree $[\QQ(i):\QQ]=2$. $\QQ(i)$ has two maps to $\CC$, by $i\mapsto i$ and $i\mapsto -i$, each of which defines a norm on $\QQ(i)$. 
\end{example}
For any number field, each conjugate of a generator defines a homomorphism to $\CC$. 
\begin{example}[$\wp$-adic Valuation]
  Let $K$ be a number field and $\Ocal_{K}$ be its ring of algebraic integers. Let $\wp$ be a prime ideal of $\Ocal_{K}$. For each $x\in K$, $(x)$ is an ideal of $\Ocal_{K}$ admitting a factorization into prime ideals. Write $v_{\wp}(x)$ as the power of $\wp$ in the ideal-factorization of $(x)$. For any constant $c>1$, let $|x|=c^{-v_{\wp}(x)}$ for $x\neq0$ and $|x|=\infty$. 
\end{example}
In the $\wp$-adic valuation, things are ``small'' when they are divisible by large $\wp$-powers. In the case of number fields $K$, these are the only types of valuations on $K$. The $\wp$-adic valuation allows us to see primes of a field without passing through any rings. 

One defines a global field by stating that all valuations on the field have some global coherence, wheras in the case of local fields, we are only looking at one prime, that is only one of the valuations of a local field. We will get to the definition of local fields soon. Let us begin with some examples. 
\begin{example}
  The real numbers $\RR$. 
\end{example}
\begin{example}
  The complex numbers $\CC$. 
\end{example}
\begin{example}
  Let $p$ be a prime. The $p$-adic numbers $\QQ_{p}$ is a local field. 
\end{example}
\begin{exercise}
  Let $K$ be a number field and $v_{\wp}$ be the valuation with respect to the prime ideal $\wp$. Let $K_{\wp}$ be the completion of $K$ with respect to the metric $v_{\wp}$. $K_{\wp}$ is a local field. 
\end{exercise}
In this way, local fields only allow us to see one prime of a global field at a time. Note that $\RR$ arises as a local field via the completion with respect to a metric, as the completion of $\QQ$ with respect to the Euclidean norm. 

There are several connections between local and global fields as statements about global fields can sometimes be reduced to statements about global fields such as via local to global principles. 
\begin{theorem}[Hasse-Minkowski]
  Let $Q$ be a quadratic form over a number field $K$. The form $Q$ has a non-trivial solution over $K$ if and only if $Q$ has non-trivial solutions over every completion of $K$. 
\end{theorem}
We will study questions like these in the context of local fields. Continuing with our discussion, we can define discrete valuation rings as in \cite[\S 1]{Serre}. We will continue developing the theory as in this text in the coming semester. 
\begin{definition}[Discrete Valuation]
  Let $K$ be a field. The field $K$ is discretely valued if there is a surjective homomorphism $v:K^{\times}\to\ZZ$ such that $v(x+y)\geq\min\{v(x),v(y)\}$ and $v(0)=\infty$. 
\end{definition}
This allows us to define a Discrete Valuation ring, or DVR. 
\begin{definition}[DVR]
  Let $K$ be a field with discrete valuation $v$. Let 
  $$A=\{x\in K|v(x)\geq0\}$$
  is a discrete valuation ring with fraction field $K$. 
\end{definition}
Let us consider the following examples. 
\begin{example}
  Let $K=\QQ$ and $p$ a prime. $A=\ZZ_{(p)}=\{\frac{r}{s}|p\nmid s; r,s\in\ZZ\}\subsetneq\QQ$ is a DVR. 
\end{example}
\begin{example}
  Let $k$ be a field and $k((t))$ be the field of Laurent series on $k$, that is for some $n_{0}>-\infty$, $\sum_{n\geq n_{0}}a_{n}t^{n}$. There is a discrete valuation on the field of Laurent series by
  $$v\left(\sum_{n\geq n_{0}}a_{n}t^{n}\right)=n_{0}.$$
  the DVR $A$ here is just the power series on $k$, $k[[t]]$.
\end{example}
\begin{definition}[Uniformizer]
  An element $\pi\in K$ is a uniformizer if $v(\pi)=1$. 
\end{definition}
Let us consider the following example.
\begin{example}
  Let $\QQ$ be a field endowed with the $p$-adic valuation. All integers with prime $p$ appearing exactly once in the prime factorization is a uniformizer. 
\end{example}
Indeed, for $x\in A\setminus\{0\}$, we can write $x=\pi^{n}u$ for some $u\in A^{\times}$. Let $\mfrak=(\pi)=\pi A$ the ideal generated by $\pi$. We can show $\mfrak=\{x\in K|v(x)\geq1\}$ and that $\mfrak$ is a maximal ideal. We define the residue field to be $A/\mfrak$. 
\section{Lecture 2 -- 13th September 2023}
Let us begin by recalling some facts about integrality. 
\begin{definition}[Integral Domain]
  A ring $A$ is an integral domain if it has no zerodivisors, that is, there do not exist $a,b\in A\setminus\{0\}$ such that $ab=0$. 
\end{definition}
\begin{definition}[Integral Over]
  Let $A,B$ be rings and $A\subseteq B$. An element $x\in B$ is integral over $A$ if it is the root of some monic polynomial with coefficients in $A$. The ring $B$ is integral over $A$ if all its elements are integral over $A$. 
\end{definition}
\begin{definition}[Integrally Closed]
  A ring $A$ is integrally closed if all elements of the fraction field $K(A)$ are integral over $A$. 
\end{definition}
Let us consider some examples. 
\begin{example}
  $\ZZ$ is integrally closed. This is the content of Gauss' lemma. 
\end{example}
\begin{example}
  Let $K$ be a number field and $\Ocal_{K}$ be its ring of integers. $\Ocal_{K}$ is integrally closed. 
\end{example}
In fact, we can show that DVRs are also integrally closed. To do so, we first show the follwing lemma. 
\begin{lemma}
  Let $A$ be a DVR with fraction field $K(A)$. If $x_{1},\dots,x_{n}\in K(A)$ are such that $v(x_{i})>v(x_{1})$ for all $i\geq 2$, then $x_{1}+\dots+x_{n}\neq0$. 
\end{lemma}
\begin{proof}
  Suppose to the contrary $x_{1}+\dots+x_{n}=0$. Equivalently $x_{2}+\dots+x_{n}=-x_{1}$ and applying the valuation we see that $v(x_{2}+\dots+x_{n})\geq\min\{v(x_{2}),\dots,v(x_{n})\}>v(x_{1})=v(-x_{1})$ a contradiction since $v(x_{2}+\dots+x_{n})\neq v(x_{1})=v(-x_{1})$. 
\end{proof}
We can now show the desired fact. 
\begin{proposition}[DVRs Integrally Closed]\label{prop:DVR integrally closed}
  A DVR $A$ is integrally closed. 
\end{proposition}
\begin{proof}
Let $x\in K(A)$ such that it satisfies the monic polynomial in $A$
$$x^{n}+a_{n-1}x^{n-1}+\dots+a_{1}x+a_{0}=0.$$
We seek to show $x\in A$, that is, $v(x)\geq0$. Suppose to the contrary that $v(x)=-m$ for some $m>0$. Applying the discrete valuation on $A$ to the polynomial, we have $v(x^{n})=-mn$ and each of the latter summands have valuations at least $-(n-1)m>-nm=v(x^{n})$ since each of the $a_{i}$ have non-negative discrete valuations. Applying the lemma above, $x$ cannot be the root of the polynomial, giving a contradiction and showing that $v(x)\geq0$ as desired. 
\end{proof}
More explicitly, we can show that the condition in \Cref{prop:DVR integrally closed} in conjunction with the condition of having a unique nonzero prime ideal to characterize DVRs. 
\begin{proposition}\label{prop:DVR iff integrally closed and unique nonzero prime}
  A Noetherian integral domain $A$ is a DVR if and only if it is integrally closed and has a unique nonzero prime ideal. 
\end{proposition}
\begin{proof}
  See \cite[Ch. 1, \S 2, Prop. 3]{Serre}. 
\end{proof}
Let us narrow our focus to Dededekind domains and understand these through the lens of DVRs. To do this, we will first have to define localization. 
\begin{definition}[Localization]
  Let $A$ be an integral domain and $K(A)$ its fraction field. Let $S\subseteq A$ be a multiplicatively closed subset of $A$ containing 1. We define the localization of $A$ away from $S$ as the set 
  $$S^{-1}A=\left\{\frac{a}{s}|a\in A, s\in S\right\}$$
  under the operations of addition and multiplication.
\end{definition}
\begin{example}
  Let $S=A\setminus\{0\}$ then $S^{-1}A=K(A)$, the fraction field of $A$. 
\end{example}
\begin{example}
  Let $S=A^{\times}$ then $S^{-1}A=A$. 
\end{example}
The prime ideals in $A$ and $S^{-1}A$ are closely related in the following way: the prime ideals in $S^{-1}A$ are those prime ideals in $A$ not intersecting $S$. 
\begin{proposition}\label{prop:localization correspondence}
  There is a bijection between prime ideals of $A$ and $S^{-1}A$ sending a prime ideal $\wp\subseteq A$ to $S^{-1}\wp=\{\frac{p}{s}|p\in\wp,s\in S\}$ and conversely sending $\wp\subseteq S^{-1}A$ to $\wp\cap A$. 
\end{proposition}
\begin{proof}
  Let $\wp$ be a prime ideal of $A$ not intersecting $S$ and consider $\frac{ab}{s_{1}s_{2}}\in S^{-1}\wp$ where $a,b\in A, s_{1},s_{2}\in S$. There is $\frac{c}{s}\in S^{-1}\wp$ such that $\frac{ab}{s_{1}s_{2}}=\frac{c}{s}$ or equivalently that for some $t\in S$, we have $t(abs-cs_{1}s_{2})=0$ where $abst = cs_{1}s_{2}t$ and $abst\in\wp$. Noting $st\in S$ with $S$ multiplicatively closed and $S\cap\wp=\emptyset$ $ab\in\wp$ since $\wp$ is prime in $A$ and one of $a,b$ is in $\wp$. So one of $\frac{a}{s_{1}},\frac{b}{s_{2}}\in S^{-1}\wp$. One can also see by definition of $S^{-1}\wp=\{\frac{p}{s}|p\in\wp,s\in S\}$ that $S^{-1}\wp\cap A=\wp$. 
  \\\\
  Let $\overline{\wp}$ be a prime ideal in $S^{-1}A$. Under the homomomorphism of rings $\varphi:A\to S^{-1}A$, we know $\varphi^{-1}(\overline{\wp})=\{a\in A|\frac{a}{1}\in\overline{\wp}\}\subseteq A$ is a prime ideal. Suppose to the contrary that $S\cap\wp\neq\emptyset$ then $1\in\overline{\wp}$ and $\overline{\wp}=S^{-1}A$ contradicting that $\overline{\wp}$ is prime in $S^{-1}A$. One can also see that $S^{-1}\wp=\{\frac{a}{s}|a\in\wp,s\in S\}=\overline{\wp}$ proving the bijective correspondence. 
\end{proof}
\begin{example}
  Let $S=A\setminus\wp$ for $\wp$ a prime ideal. We write $A_{\wp}=S^{-1}A$ the localization of $A$ away from $\wp$. The ideals of $A_{\wp}$ are exactly those primes contained in $\wp$ so $A_{\wp}$ has a unique maximal ideal, that is, $A_{\wp}$ is a local ring. 
\end{example}
Introducing some notation, let $I\subseteq A$ be an ideal. We denote $I_{\wp}\subseteq A_{\wp}$ the ideal in $A_{\wp}$ generated by $I$. We will soon prove some equivalent conditions for a Noetherian integral domain to be a Dedekind domain. To do so, we require the following lemma. 
\begin{lemma}\label{lem:intersection of prime ideals}
  Let $A$ be an integral domain. We have an equality $\cap_{\wp\text{ prime}}A_{\wp}=A$. 
\end{lemma}
\begin{proof}
  Let $x\in\cap_{\wp\text{ prime}}A_{\wp}$ and write $x=\frac{a}{b}$. Consider the ideal $\afrak\subseteq A$ by $\afrak=\{c\in A|ca\in(b)\subseteq A\}$ which can be thought of as the ideal of denominators of $x$. For every prime $\wp$, $\afrak\not\subseteq\wp$ as $x\in A_{\wp}$ but every proper ideal is contained in some maximal ideal so we have $\afrak=A$, showing that $x\in A$. 
\end{proof}
We can now show the following theorem that we will use to characterize Dedekind domains. 
\begin{theorem}\label{thm: dedekind TFAE}
  Let $A$ be a Noetherian integral domain. The following are equivalent:
  \begin{enumerate}[label=(\roman*)]
    \item For every nonzero prime ideal $\wp\subseteq A$, $A_{\wp}$ is a DVR. 
    \item $A$ is integrally closed and of Krull dimension at most 1. 
  \end{enumerate}
\end{theorem}
This is \cite[Ch. 1, \S 1, Prop. 4]{Serre}. 
\begin{proof}
  (i)$\Longrightarrow$(ii) Suppose $\wp\subseteq A$ is a prime ideal and $\mfrak$ the maximal ideal containing $\wp$. The local ring $A_{\mfrak}$ contains the prime ideal $\wp_{\mfrak}$ by \Cref{prop:localization correspondence}. Since $\mfrak$ is prime, $A_{\mfrak}$ is a DVR by assumption and given that the sole prime ideal in a DVR is $(\pi)$ for $\pi$ some uniformizer, we have either $\wp_{\mfrak}=(0)$ or $\wp_{\mfrak}=(\pi)$ showing that $A$ is of Krull dimension 1. To see that $A$ is integrally closed, note that for some $x\in K(A)$ be integral over $A$. We want to show that $x\in A$ but $x\in A_{\wp}$ for all primes $\wp$ so by \Cref{lem:intersection of prime ideals}, $x\in A$ showing that $A$ is integrally closed. 
  \\\\
  (ii)$\Longrightarrow$(i) Suppose that $A$ is integrally closed and of Krull dimension at most 1. Let $\wp\subseteq A$ be a nonzero prime ideal and consider the localization $S^{-1}A$ for $S=A\setminus\wp$. The ring $S^{-1}A$ has a unique nonzero prime ideal $S^{-1}\wp$. By \Cref{prop:DVR iff integrally closed and unique nonzero prime}, it suffices to show that $S^{-1}A$ is integrally closed. Let $x\in K(S^{-1}A)=K(A)$ be integral over $S^{-1}A$ so it satisfies
  \begin{align}\label{eqn: dedekind domain TFAE}
    x^{n}+b_{n-1}x+\dots+b_{1}x+b_{0} &= 0
  \end{align}
  for some $b_{i}\in S^{-1}A$. Writing each $b_{i}=\frac{a_{i}}{s_{i}}$ for $a_{i}\in A, s_{i}\in S$, we can take $s=\prod_{i=0}^{n-1}s_{i}$ and multiply (\ref{eqn: dedekind domain TFAE}) by $s^{n}$ to yield 
  \small
  \begin{align*}
    0 &= s^{n}x^{n}+a_{n-1}\left(s_{n-1}^{n-1}\prod_{0\leq i\leq n-2}s_{i}^{n}\right)x^{n-1}+\dots+a_{1}\left(s_{1}^{n-1}\prod_{0\leq i\leq n-1, i\neq 1}s_{i}^{n}\right)x+a_{0}\left(s_{0}^{n-1}\prod_{1\leq i\leq 0}s_{i}^{n}\right) \\
    &= (sn)^{n} + a_{n-1}\left(\prod_{0\leq i\leq n-2}s_{i}\right)(sx)^{n-1}+\dots+a_{1}\left(s_{1}^{n-2}\prod_{0\leq i\leq n-1, i\neq 1}s_{i}^{n-1}\right)(sx)+a_{0}\left(s_{0}^{n-1}\prod_{1\leq i\leq 0}s_{i}^{n}\right)
  \end{align*}
  \normalsize
  wherein $sx\in A$ and each 
  \small
  $$a_{n-1}\left(\prod_{0\leq i\leq n-2}s_{i}\right), \dots, a_{1}\left(s_{1}^{n-2}\prod_{0\leq i\leq n-1, i\neq 1}s_{i}^{n-1}\right), a_{0}\left(s_{0}^{n-1}\prod_{1\leq i\leq 0}s_{i}^{n}\right)\in A.$$
  \normalsize
  so by $A$ being integrally closed we conclude $sx\in A$ and thus $x\in S^{-1}A$ as desired. 
\end{proof}
We use these equivalent conditions to define a Dedekind domain. 
\begin{definition}[Dedekind Domain]
  A Dedekind domain is a Noetherian integral domain $A$ that satisfies any one of the equivalent conditions in \Cref{thm: dedekind TFAE}. 
\end{definition}
Some examples of Dedekind domains are as follows. 
\begin{example}
  $\ZZ, \Ocal_{K}$ for $K$ a number field, $\FF_{q}[t]$, and $\CC[t]$ are all Dedekind domains. 
\end{example}
\begin{example}
  Any PID is a Dedekind domain. Using the fact that PIDs are UFDs, one can yield a valuation by considering factorizations of the ideal generated by an element into prime ideals. 
\end{example}
One can also show that the property of being a Dedekind domain is preserved under localization. 
\begin{proposition}
  If $A$ is a Dedekind domain and $S$ a multiplicative subset of $A$, the localization $S^{-1}A$ is a Dedekind domain. 
\end{proposition}
\begin{proof}
  We show condition (b) in \Cref{thm: dedekind TFAE}. By hypothesis $A$ is a Dedekind domain so it is integrally closed and of Krull dimension at most 1. Since prime ideals of $S^{-1}A$ are those of $A$ not meeting $S$, $S^{-1}A$ is also of Krull dimension at most 1. One then applies the proof in the latter part of \Cref{thm: dedekind TFAE} to show integral closure. 
\end{proof}
Let $A$ be an integral domain and $K(A)$ its field of fractions. For $x\in K(A)$, we want to assign a valuation to $A$ with respect to some prime ideal $\wp$ of $A$. We make sense of this in terms of fractional ideals which provide a notion of multiplicative inverse for ideals. 
\begin{definition}[Fractional Ideal]
  Let $A$ be an integral domain with fraction field $K(A)$. A fractional ideal of $A$ is a finitely generated $A$-submodule of $K(A)$. 
\end{definition}
Given two $A$-submodules of $K(A)$, we can multiply them 
$$IJ=\left\{\sum_{i=1}^{n}a_{i}b_{i}|a_{i}\in I,b_{i}\in J\right\}.$$
\begin{definition}[Invertible]
  An ideal $I$ is invertible if there exists $J$ such that $IJ=A$. 
\end{definition}
In the case of Dedekind domainsm we can show that every ideal is invertible. This is a key part of the proof of the unique factorization property in Dedekind domains, and can in fact be used to define Dedekind domains themselves. 
\begin{proposition}[Dedekind Ideals are Invertible]
  Let $A$ be a Dedekind domain. Every nonzero fractional ideal is invertible. 
\end{proposition}
\begin{proof}
  Let $A$ be a Dedekind domain and $I\subseteq A$ a fractional ideal. Let $J=\{x\in K|xI\subseteq A\}$. We have $IJ\subseteq A$. Suppose to the contrary that $IJ\subsetneq A$. The ideal $IJ$ is contained in some maximal ideal $\mfrak$ so localizing at $\mfrak$ we have $J_{\mfrak}=\{x\in K|x I_{\mfrak}\in A_{\mfrak}\}$, that is, the inverse of $I_{\mfrak}$ as fractional ideals of $A_{\mfrak}$. In particular, observe $(IJ)_{\mfrak}=I_{\mfrak}J_{\mfrak}=A_{\mfrak}$, a contradiction as $IJ\subseteq\mfrak$. 
\end{proof}
We apply this to the factorization of ideals. 
\begin{theorem}[Factorization in Dedekind Domains]
  Let $A$ be a Dedekind domain. Every fractional ideal $I$ can be written uniquely as 
  $$I=\prod_{\wp_{i}\text{ prime}}\wp_{i}^{e_{i}}$$
  only finitely many $e_{i}$ nonzero. 
\end{theorem}
The proof of uniqueness was given in class. For existence, we adapt the proof from \cite[\S 3, Thm. 14]{Marcus}.
\begin{proof}
  (Existence) Suppose to the contrary that there is some collection of ideals not admitting a decomposition into prime dieals as above. Let $\qfrak$ be the maximal among such ideals not admitting a decomposition into primes where $\qfrak\neq A$ as $A$ is the empty product. Let $\pfrak$ be a prime ideal of $A$ containing $\qfrak$ and note that $\pfrak=\qfrak\afrak$ for some ideal $\afrak\subseteq A$ since $\qfrak\subseteq\pfrak$ implies $\pfrak|\qfrak$. But $\afrak$ contains $\qfrak$ so $\qfrak|\afrak$ and in fact $\qfrak\subsetneq\afrak$ as otherwise $\afrak=\qfrak$ and $\qfrak=\pfrak\qfrak$ implying $A=\pfrak$ a contradiction as $\pfrak$ was taken to be prime and hence proper. Using $\qfrak\subsetneq\afrak$ we have $\afrak$ is a product of primes and so is $\qfrak$, a contradiction. 
  \\\\
  (Uniqueness) Let $I\subseteq A$ be an ideal admitting a decomposition by 
  $$I=\prod_{\wp_{i}\text{ prime}}\wp_{i}^{e_{i}}.$$
  For each $\wp_{i}$, $I_{\wp_{i}}$ is an ideal in the DVR $A_{\wp_{i}}$ and $e_{i}=v_{\wp_{i}}(x)$ for $x$ a generator of $I_{\wp_{i}}$ in $A_{\wp_{i}}$.
\end{proof}
Returning to integrality, we prove the following statement that we will later use. 
\begin{proposition}\label{prop:integral over iff finite generation as a module}
  Let $A\subseteq B$ be rings. The elements $b_{1},\dots,b_{n}\in B$ are integral over $A$ if and only if $A[b_{1},\dots,b_{n}]$ is finitely generated as an $A$-module. 
\end{proposition}
Note that $A[b_{1},\dots,b_{n}]$ is finitely generated as an $A$-algebra, we are asking for something stronger. 
\begin{proof}
  $(\Longrightarrow)$ By induction, it suffices to show that if $b\in B$ is integral over $A$ the ring $A[b]$ is finitely generated as an $A$-module. Suppose $b\in B$ is integral over $A$ so 
  $$b^{n}+a_{n-1}b^{n-1}+\dots+a_{1}b+a_{0}=0$$
  for some $a_{0}, a_{1},\dots, a_{n-1}\in A$. Thus $A[b]$ is generated by $1, b, \dots,b^{n-1}$ as an $A$-module, that is, finitely generated as an $A$-module. 
  \\\\
  $(\Longleftarrow)$ Suppose $w_{1},\dots,w_{m}$ generate $A[b_{1},\dots,b_{n}]$ as an $A$-module. For $b\in A[b_{1},\dots,b_{n}]$ and any $w_{i}$ we can write 
  $$bw_{i}=\sum_{j=1}^{m}a_{ij}w_{j}$$
  with $a_{ij}\in A$. We use this $A$-linear expression to construct an integral dependence polynomial. Let 
  $$M=\begin{bmatrix}
    b-a_{11} & -a_{12} & \dots & -a_{1m} \\
    -a_{21} & b-a_{22} & \dots & a_{2m} \\
    \vdots & \vdots & \ddots & \vdots \\
    -a_{m1} & -a_{m2} & \dots & b-a_{mm}
  \end{bmatrix}.$$
  Observe that 
  $$M\begin{bmatrix}
    w_{1} \\ \vdots \\ w_{m}
  \end{bmatrix}=0_{m}$$
  so for $M^{*}$ the adjoint matrix satisfying $M^{*}M=\det M I_{m}$ we can consider the equation 
  $$M^{*}M\begin{bmatrix} w_{1}\\ \vdots\\ w_{m}\end{bmatrix}=\det M\begin{bmatrix}
    w_{1} \\ \vdots \\ w_{m}
  \end{bmatrix}=0_{m}$$
  as $1$ can be written as a linear cobmination of the $w_{i}$s. Thus $\det M$ is a monic in $A$ satisfied by $b\in A[b_{1},\dots,b_{n}]$ showing that $A[b_{1},\dots,b_{n}]$ is integral over $A$. 
\end{proof}
\section{Lecture 3 -- 18th September 2023}
Let us recall \Cref{prop:integral over iff finite generation as a module} with which we ended the previous lecture. 
\begin{proposition}[=\Cref{prop:integral over iff finite generation as a module}]
  Let $A\subseteq B$ be rings. The elements $b_{1},\dots,b_{n}\in B$ are integral over $A$ if and only if $A[b_{1},\dots,b_{n}]$ is finitely generated as an $A$-module. 
\end{proposition}
One can deduce the following corollaries. 
\begin{corollary}\label{corr: integral elements are subring}
  Let $A\subseteq B$ be rings. The elements of $B$ integral over $A$ form a subring of $B$. 
\end{corollary}
% Maybe edit this. 
\begin{proof}
  If $b_{1},b_{2}\in B$ are integral over $A$ then $b_{1}b_{2}, b_{1}+b_{2}\in A[b_{1},b_{2}]$. 
\end{proof}
\begin{corollary}
  Let $A\subseteq B\subseteq C$ be a tower of rings. If $B$ is integral over $A$ and $C$ is integral over $B$ then $C$ is integral over $A$. 
\end{corollary}
% Maybe edit this. 
\begin{proof}
  If $c_{i}$ generate $C$ over $B$ and the $b_{j}$ generate $B$ over $A$ then the $c_{i}b_{j}$ generate $C$ over $A$. 
\end{proof}
We now turn to extensions of Dedekind domains. More precisely, let $A$ be a Dedekind domain and $K$ its fraction field. Let $B$ be the integral closure of $A$ in $L$. 
$$% https://q.uiver.app/#q=WzAsNCxbMCwwLCJCIl0sWzAsMSwiQSJdLFsyLDAsIkwiXSxbMiwxLCJLIl0sWzEsMywiIiwyLHsic3R5bGUiOnsidGFpbCI6eyJuYW1lIjoiaG9vayIsInNpZGUiOiJ0b3AifX19XSxbMCwyLCIiLDIseyJzdHlsZSI6eyJ0YWlsIjp7Im5hbWUiOiJob29rIiwic2lkZSI6InRvcCJ9fX1dLFswLDFdLFsyLDNdXQ==
\begin{tikzcd}
	B && L \\
	A && K
	\arrow[hook, from=2-1, to=2-3]
	\arrow[hook, from=1-1, to=1-3]
	\arrow[from=1-1, to=2-1]
	\arrow[from=1-3, to=2-3]
\end{tikzcd}$$
Let us consdier some examples from the number field and function field case. 
\begin{example}
  For a cubic extension of fields $\QQ(\sqrt[3]{2})/\QQ$, both $\ZZ$ and $\ZZ[\sqrt[3]{2}]$ are Dedekind domains. 
  $$% https://q.uiver.app/#q=WzAsNCxbMCwwLCJaWltcXHNxcnRbM117Mn1dIl0sWzAsMSwiXFxaWiJdLFsyLDAsIlxcUVEoXFxzcXJ0WzNdezJ9KSJdLFsyLDEsIlxcUVEiXSxbMSwzLCIiLDIseyJzdHlsZSI6eyJ0YWlsIjp7Im5hbWUiOiJob29rIiwic2lkZSI6InRvcCJ9fX1dLFswLDIsIiIsMix7InN0eWxlIjp7InRhaWwiOnsibmFtZSI6Imhvb2siLCJzaWRlIjoidG9wIn19fV0sWzAsMV0sWzIsM11d
  \begin{tikzcd}
    {\ZZ[\sqrt[3]{2}]} && {\QQ(\sqrt[3]{2})} \\
    \ZZ && \QQ
    \arrow[hook, from=2-1, to=2-3]
    \arrow[hook, from=1-1, to=1-3]
    \arrow[from=1-1, to=2-1]
    \arrow[from=1-3, to=2-3]
  \end{tikzcd}$$
\end{example}
\begin{example}
  Fix some finite field $\FF_{q}$ of characteristic not 2. Let $A=\FF_{q}[t]$ with field of fractions $K=\FF_{q}(t)$. Let 
  $$L=\FF_{q}(t)[\sqrt{t^{3}+t+1}]=\FF_{q}(t)[s]/(s^{2}-t^{3}-t-1).$$
  One can verify that $[L:K]=2$ where $K$ are the functions on $\PP^{1}_{\FF_{q}}$ and $L$ are the functions on the elliptic curve $s^{2}=t^{3}+t+1$ (more often written $y^{2}=x^{3}+x+1$) and the extension of fields $L/K$ corresponds to the 2-1 map $E\to\PP^{1}_{\FF_{q}}$ from the elliptic curve to the projective line. The integral closure of $L$ in $A=\FF_{q}[t]$ is $B=\FF_{q}[t][\sqrt{t^{3}+t+1}]$. 
\end{example}
\begin{remark}
  One could also take $A=\FF_{q}[\frac{1}{t}]$ and the same extensions $L/K$. But now $\sqrt{t^{3}+t+1}$ is no longer integral over $A$. 
\end{remark}
\begin{proposition}
  If $A$ is a Dedekind domain with fraction field $K$ and $B$ the integral closure of $A$ in $L$ for $L/K$ a finite extention of fields, the ring $B$ is an integral domain with fraction field $L$. 
\end{proposition}
\begin{proof}
  The integral closure of $A$ in $L$ is a subring by \Cref{corr: integral elements are subring}. It remains to show that $B$ has fraction field $L$. 
\end{proof}
We want to show a stronger fact, that $B$ is in fact itself a Dedekind domain. This parallels what we have seen before. For $K$ a number field and $L/K$ a finite extension of fields, $\Ocal_{L}$ is also a Dedekind domain. We are seeking a proof that generalizes appropriately to non-separable extensions and the function field case. In this setting, we lose nice properties such as separability. But first, let us return to the simpler case of separable extensions. 
\\\\
To consider this case, we recall some notions from our first course in algebraic number theory. For an extension of fields $L/K$, multiplication by an element $x\in L$ gives a linear transformation of $K$-vector spaces $m_{x}:L\to L$ by $\alpha\mapsto x\alpha$. The linear transformation admits a trace which we denote $\Tr_{L/K}(x)=\mathrm{tr}(m_{x})$ and a norm $\Nm_{L/K}(x)=\det(m_{x})$. If further $L/K$ is a Galois extension, we can compute $\Tr_{L/K}(x)=\sum_{\sigma\in\Gal(L/K)}\sigma(x)$ and $\Nm_{L/K}(x)=\prod_{\sigma\in\Gal(L/K)}\sigma(x)$. Finally, recall that the $K$-bilinear form $\langle x,y\rangle=\Tr_{L/K}(xy)$ is nondegenerate if and only if $L/K$ is separable -- a bilinear form is nondegenerate if $\langle x,y\rangle=0$ for all $y$ then $x=0$. We can now prove the following. 
\begin{proposition}
  Let $A$ be a Dedekind domain with fraction field $K$ and $B$ the integral closure of $A$ in $L$ for $L/K$ a finite extention of fields. If further $L/K$ is separable, the ring $B$ is a finitely generated $A$-module (i.e. satisfies Serre's ``Hypothesis F''). 
\end{proposition}
This is \cite[Ch. 1, \S 4, Prop. 8]{Serre}. 
\begin{proof}
  For $x\in B$ we know it is integral over $A$ by hypothesis so $\Tr_{L/K}(x)\in A$ as it is an integral multiple of the coefficient of $x$ by \cite[\href{https://stacks.math.columbia.edu/tag/0BIH}{0BIH}]{stacks-project}. Let $\{e_{i}\}$ be a basis of $L$ over $K$. Without loss of generality, we can take $\{e_{i}\}\subseteq B$ by clearing denominators. Let $V$ be the free $A$-module spanned by the $\{e_{i}\}$. Using the trace map, we consider the dual module 
  $$V^{*}=\left\{x\in L|\Tr_{L/K}(xy)\in A, \forall y\in V\right\}.$$
  We have $V\subseteq B$ by inspection as the $\{e_{i}\}$ were chosen in $B$ and $B^{*}\subseteq V^{*}$ where $V^{*}$ is the free module spanned by the basis dual to $\{e_{i}\}$ with respect to the trace bilinear form the existence of which is given by separability. In particular, $B$ is an $A$-submodule of $V^{*}$ which was finitely generated by the finite dual basis. 
\end{proof}
To show that integral closures of Dedekind domains are Dedekind domains, however, will require some lemmata, some of which was assigned as homework. 
\begin{lemma}\label{lem: equality of primes in B}
  If $\wp\subseteq\wp'$ are prime ideals of $B$ and $\wp\cap A=\wp'\cap A$ then $\wp=\wp'$. 
\end{lemma}
\begin{proof}
  Consider the quotient $B/\wp$ and for $x\in\wp'\setminus\wp$ its image $\overline{x}\in B/\wp$. Quotients are homomorphisms of rings so
  $$\overline{x}^{n}+\overline{a_{n-1}}\cdot\overline{x}^{n-1}+\dots+\overline{a_{1}}\cdot\overline{x}+\overline{a_{0}}=0$$
  and rearranging we see that $a_{0}\in(\overline{x})\subseteq\overline{A}$, that is, $a_{0}\in\overline{\wp'\cap A}$ but not $\overline{\wp\cap A}$ a contradiction. 
\end{proof}
\begin{lemma}\label{lem:descending chain stabilizes}
  In a 0-dimensional Noetherian ring, any descending chain of ideals stabilizes. 
\end{lemma}
\begin{lemma}\label{lem: ring B0}
  Let $A$ be a Dedekind domain with fraction field $K$ and $B$ the integral closure of $A$ in $L$ for $L/K$ a finite extension of fields. Let $w_{1},\dots,w_{n}\in L$ be an $L$-basis fully contained in $B$ and denote $B_{0}=A[w_{1},\dots,w_{n}]$. If $a\in A$ then $B_{0}/aB_{0}$ is a 0-dimensional Noetherian ring. 
\end{lemma}
Granting the lemmata, we can prove the theorem. 
\begin{theorem}\label{thm: integral closure is Dedekind}
  Let $A$ be a Dedekind domain with fraction field $K$ and $B$ the integral closure of $A$ in $L$ for $L/K$ a finite extension of fields. The ring $B$ is a Dedekind domain. 
\end{theorem}
\begin{proof}
  We seek to show that $B$ is Noetherian, integrally closed and of Krull dimension 1. By construction, $B$ is integrally closed. Suppose to the contrary that $B$ is of Krull dimension more than 1 and we have a strict chain of prime ideals $\wp_{1}\subsetneq\wp_{2}\subsetneq_{3}$ of $B$. Contraposing \Cref{lem: equality of primes in B} and intersecting down to $A$ we have $\wp_{1}\cap A\subsetneq\wp_{2}\cap A\subsetneq\wp_{3}\cap A$, a contradiction as $A$ was Dedekind by assumption and of Krull dimension at most 1. \\\\
  It remains to show $B$ is Noetherian. Let $w_{1},\dots,w_{n}$ be a $K$-basis of $L$ contained in $B$ with each $w_{i}$ integral over $A$ and let $B_{0}=A[w_{1},\dots,w_{n}]$. By \Cref{prop:integral over iff finite generation as a module}, $B_{0}$ is a finitely generated $A$-module and therefore Noetherian. To show $B$ is itself Noetherian, we will show that any ideal $I\subseteq B$ is a finitely generated $B$-module for nonzero ideals $I$. We claim that for $a\in I\cap A$, $B/aB$ is a finitely generated $B$-module. Granting \Cref{lem: ring B0}, consider the decreasing sequence of ideals in $B_{0}$ given by $(a^{m}B\cap B_{0},aB_{0})$ for $m\geq 1$ which correspond to a descending sequence $\overline{(a^{m}B\cap B_{0})}$ in $B_{0}/aB_{0}$. By \Cref{lem: ring B0}, they stabilize at some $m\geq n$. For $\beta\in B$, we want to show that $\beta\in a^{-n}B_{0}+aB$. Take $h$ minimal such that $\beta$ can be expressed as $a^{-h}B_{0}+aB$. If $h\leq n$ we are done. Otherwise suppose to the contrary that $h>n$. Let $\beta=\frac{u}{a^{h}}+a\widetilde{u}$ for $u\in B_{0},\widetilde{u}\in B$. Rearranging this expression, we have $a^{h}(\beta-a\widetilde{u})$ so $u\in a^{h}B\cap B_{0}$ which lies in $I_{h}$ but $h>n$ contradicting the stabilization if ideals so $u\in I_{h}=I_{h-1}$ so $a^{h-1}\widetilde{u'}+au'$ so $u'\in B_{0},\widetilde{u'}\in B$ so $\beta=\frac{u'}{a^{h-1}}+a(\widetilde{u}+\widetilde{u'})$ contradicting minimality of $h$. So $B/aB\subseteq(a^{-n}B_{0}\cap aB)/aB$ and $B/aB$ is a finitely generated $B_{0}$-module. Then $B/aB$ is a finitely generated $A$ module and for an ideal $I$ containing $aB$ of $B$, $I/aB$ is a finitely genereated $A$-module in $B/aB$ with $A$ Noetherian, so $I$ is a finitely generated $B$-module. 
\end{proof}
\section{Lecture 4 -- 20th September 2023}
We continue our discussion of extensions of Dedekind domains. Let us recall the setup. Let $A$ be a Dedekind domain with fraction field $K$ and $B$ the integral closure of $A$ in $L$, where $L/K$ is a finite extension of fields. By \Cref{thm: integral closure is Dedekind}, $B$ is a Dedekind domain. Let $\pfrak$ be a nonzero prime ideal of $A$. Write $\wp=\pfrak B$ be its ``lift'' in $B$. Since $B$ is Dedekind, $\wp$ admits a unique factorization into prime ideals $\prod_{i}\wp_{i}^{e_{i}}$ finitely many $e_{i}$ nonzero. In such a setting, we write $\wp_{i}|\pfrak$. 
\begin{definition}[Ramifies]
  Let $\pfrak\subseteq A$ be an ideal and $\pfrak B=\prod_{i}\wp_{i}^{e_{i}}$ with finitely many $e_{i}$ nonzero. If $e_{i}\geq1$, the ideal $\wp_{i}\subseteq B$ ramifies with ramification index $e_{i}$ in $L/K$. Similarly if $e_{i}\geq 1$ and $\wp_{i}|\pfrak$ for $\pfrak\subseteq A$ then $\pfrak$ ramifies in $A$. 
\end{definition}
\begin{proposition}
  If $\pfrak\subseteq A$ is an ideal and $\pfrak B=\prod_{i}\wp_{i}^{e_{i}}$ with finitely many $e_{i}$ nonzero then $\wp_{i}\cap A=\pfrak$. 
\end{proposition}
\begin{proof}
  We have an injective map of rings $A/(\wp_{i}\cap A)\to B/\wp_{i}$ where $B/\wp_{i}$ is integral as $\wp_{i}$ is prime so we have $\pfrak\subseteq\pfrak B\cap A\subseteq\wp_{i}\cap A$ but since $\pfrak$ is nonzero, $\pfrak=\wp_{i}\cap A$ as desired. 
\end{proof}
\begin{definition}[Inertia Degree]
  Let $\pfrak\subseteq A$ be a nonzero ideal and $\pfrak B=\prod_{i}\wp_{i}^{e_{i}}$ with finitely many $e_{i}$ nonzero with $\pfrak,\wp_{i}$ maximal. We define the inertia degree of $\wp_{i}$ in $L/K$ by $f_{\wp_{i}}=[B/\wp_{i}:A/\pfrak]$.  
\end{definition}
\begin{definition}[Split Completely]
  Let $\pfrak\subseteq A$ be a nonzero ideal and $\pfrak B=\prod_{i}\wp_{i}^{e_{i}}$ with finitely many $e_{i}$ nonzero with $\pfrak,\wp_{i}$ maximal. The ideal $\pfrak$ is split completely if $e_{i}=f_{\wp_{i}}=1$ for all $\wp_{i}$. 
\end{definition}
Inertia and ramification degrees are connected in the following way:
\begin{proposition}
  Let $\pfrak\subseteq A$ be a nonzero ideal and $\pfrak B=\prod_{i}\wp_{i}^{e_{i}}$ with finitely many $e_{i}$ nonzero. Under the conditions of Serre's Hypothesis F, where $B$ is a finitely generated $A$-module, $[L:K]=\sum_{\wp_{i}}e_{i}f_{\wp_{i}}$. 
\end{proposition}
\begin{proof}
  This is \cite[Ch. 1, \S 4, Proposition 10]{Serre}. 
\end{proof}
Note that without Serre's Hypothesis F, we only have the inequality $[L:K]\geq\sum_{\wp_{i}}e_{i}f_{\wp_{i}}$. \\\\
For $A$ a Dedekind domain, the set of fractional ideals of $A$ forms a group in a natural way which we denote $I_{A}$. For $B$ the integral closure of $A$, we can define a map $I_{B}\to I_{A}$ using the relative ideal norm as follows. 
\begin{definition}[Relative Ideal Norm]
  Let $A\subseteq B$ be Dedekind domains. The relative ideal norm $\Nm_{B/A}:I_{B}\to I_{A}$ is defined by $\wp\mapsto(\wp\cap A)^{f_{\wp}}$. 
\end{definition}
The relative ideal norm and the norm on elements are related in the following way. 
\begin{proposition}
  If $x\in L$ then $\Nm_{B/A}(xB)=\Nm_{L/K}(x)A$. 
\end{proposition}
If further $L/K$ is a Galois extension, there is an action by $\Gal(L/K)$ on the primes $\wp_{i}\subseteq B$ over $\pfrak\subseteq A$. 
\begin{proposition}\label{prop: transitive Galois EDD}
  Let $A$ be a Dedekind domain with fraction field $K$, $L/K$ a Galois extension of field with $B$ the integral closure of $A$ in $L$, and $\pfrak\subseteq A$ nonzero. The Galois group $\Gal(L/K)$ acts transitively on $\wp\subseteq B$ for $\wp|\pfrak$. 
\end{proposition}
\begin{proof}
  Let $\wp|\pfrak, \wp'|\pfrak$ and suppose $\wp,\wp'$ do not lie in the same Galois orbit. Applying the Chinese remainder theorem, we can find $b\in\wp$ scu that $b\notin\sigma(\wp')$, that is, $b$ does not lie in any Galois orbit of $\wp'$. We thus have $\sigma^{-1}(b)\notin\wp$ for any $\sigma\in\Gal(L/K)$. So on one hand $\Nm_{L/K}(b)\in\wp$ and computing by $\Nm_{L/K}(b)=\prod_{\sigma\in\Gal(L/K)}\sigma^{-1}(b)$ is not in $\wp'$ contradicting $\wp'\cap A=\wp\cap A$. 
\end{proof}
One then deduces the following corollary. 
\begin{corollary}
  In the setting of \Cref{prop: transitive Galois EDD}, $e_{i}=e_{j}$ and $f_{\wp_{i}}=f_{\wp_{j}}$. 
\end{corollary}
We now introduce two important subgroups of the Galois group of $L/K$, the decomposition group and the inertia group. These groups allow us to better understand the Galois group as a whole. 
\begin{definition}[Decomposition Group]
  Let $\wp|\pfrak$ for $\pfrak\subseteq A,\wp\subseteq B$ nonzero. The decomposition group of $\wp$ denoted $D_{\wp}$ is the subgroup of the Galois group $\Gal(L/K)$ stabilizing $\wp$. 
\end{definition}
By \Cref{prop: transitive Galois EDD} and our first course in the theory of groups $D_{\wp}$ and $D_{\wp'}$ are conjugate subgroups of $\Gal(L/K)$. Conversely, if $D_{\wp}, D_{\wp'}$ are conjugate subgroups of $\Gal(L/K)$, then they are decomposition groups for ideals $\wp|\pfrak, \wp'|\pfrak$. So each $\pfrak\subseteq A$ determines a conjugacy class of subgroups of $\Gal(L/K)$ which correspond to decomposition groups of ideals $\wp|\pfrak$. By the orbit-stabilizer theorem $|D|=ef=\frac{n}{r}$ where $n=[L:K]$ and $r$ is the number of primes of $B$ lying over $\pfrak\subseteq A$. Note that given a decomposition group $D$, we can consider the subfield of $L$ fixed by $D$ which we denote $K_{D}$ where $K\subseteq K_{D}\subseteq L$. One applies the orbit-stabilizer theorem once again to see that $[K_{D}:K]=r, [L:K_{D}]=ef$. Furthermore, one can show that all factorizations of $\pfrak$ into distinct primes occurs in $K_{D}$. 
\\\\
If $A/\pfrak$ is finite (which holds for our cases of interest), the decomposition group $D$ acts on $B/\wp$ fixing $A/\pfrak$ inducing a map $\phi:D\to\Gal((B/\wp)/(A/\pfrak))$. We define the kernel of this map to be the inertia group. 
\begin{definition}[Inertia Group]
  Let $\wp|\pfrak$ for $\pfrak\subseteq A,\wp\subseteq B$ nonzero with$A/\pfrak$ finite and $D_{\wp}$ the decomposition group of $\wp$. The inertia group of $\wp$ denoted $T_{\wp}$ is the kernel of the map $\phi:D\to\Gal((B/\wp)/(A/\pfrak))$. 
\end{definition}
The inertia group is the set of elements of $D$ that fix $B/\wp$ pointwise. We can think of it as the set of things that $B/\wp$ cannot see. As in the case of the decomposition group, we can define a field $K_{T}$ the subfield of $L$ fixed by the inertia group $T$. 
\begin{proposition}\label{prop:D/P is Galois group}
  Let $\wp|\pfrak$ for $\pfrak\subseteq A,\wp\subseteq B$ nonzero with $A/\pfrak$ finite, $D_{\wp}$ the decomposition group of $\wp$, and $T_{\wp}$ the inertia group of $\wp$. The map $D_{\wp}/T_{\wp}\to\Gal((B/\wp)/(A/\pfrak))$ is an isomorphism. 
\end{proposition}
\begin{proof}[Proof Outline]
  By the first isomorphism theorem for groups, it suffices to show that $\phi:D\to\Gal((B/\wp)/(A/\pfrak))$ is surjective. We can find an element $\sigma\in\Gal(L/K)$ mapping an element to each conjugate by comparing minimal polynomials of elements in $B$ and $B/\wp$. 
\end{proof}
From \Cref{prop:D/P is Galois group}, we see that $|T|=e$. 
\begin{remark}
  The inertia group $T$ is a misleading name as it controls the ramification of primes. This is the first example of the higher ramification groups that we will see later in the course. Moreover, under the assumptions above, we are considering a finite field extension over a finite field, which is well understood. 
\end{remark}
\begin{proposition}
  Let $L/K$ be a Galois extension. If $L/K$ is unramified at $\wp\subseteq B$, then $D=\Gal((B/\wp)/(A/\pfrak))$ and $D$ is generated by the Frobenius element $\Frob_{\wp}\in D$ where $\Frob_{\wp}:B/\wp\to B/\wp$ by $x\mapsto x^{|A/\pfrak|}$. 
\end{proposition}
We will often use the actions of these groups on $L/K$ and $B/\wp,A/\pfrak$ in our study of local fields. 
\begin{example}
  Let $A=\ZZ, K=\QQ, L=\QQ(\zeta_{n})$ where $\zeta_{n}$ is a primitive $n$-th root of unity. Recall here that $\Gal(L/K)\cong(\ZZ/n\ZZ)^{\times}$ where the isomorphism is by $u\mapsto \zeta_{n}^{u}$. One can check using discriminants that if $p\nmid n$ a positive prime, it is unramified and $\Frob_{p}$ is the image of $p$ in the map $(\ZZ/n\ZZ)^{\times}\to\Gal(L/K)$. Indeed, infinitely many $p$ in each least residue class $(\ZZ/n\ZZ)^{\times}$ $\Frob_{p}$ generates the Galois group. 
\end{example}
\begin{remark}
  The order of the Frobenius element $\Frob_{\wp}$ is $f_{\wp}$. 
\end{remark}
In the setting of non-Galois extensions, we can take the Galois closure and use these groups in the ways we described above. 
\begin{proposition}[={\cite[Ch. 1, \S7, Prop. 22]{Serre}}]
  Suppose we have a tower of field with their integral closures 
  $$% https://q.uiver.app/#q=WzAsNixbMCwwLCJDIl0sWzAsMSwiQiJdLFswLDIsIkEiXSxbMSwyLCJLIl0sWzEsMSwiTCJdLFsxLDAsIk0iXSxbMCw1LCIiLDAseyJzdHlsZSI6eyJ0YWlsIjp7Im5hbWUiOiJob29rIiwic2lkZSI6InRvcCJ9fX1dLFsxLDQsIiIsMCx7InN0eWxlIjp7InRhaWwiOnsibmFtZSI6Imhvb2siLCJzaWRlIjoidG9wIn19fV0sWzIsMywiIiwwLHsic3R5bGUiOnsidGFpbCI6eyJuYW1lIjoiaG9vayIsInNpZGUiOiJ0b3AifX19XSxbMSwwLCIiLDEseyJzdHlsZSI6eyJ0YWlsIjp7Im5hbWUiOiJob29rIiwic2lkZSI6InRvcCJ9fX1dLFsyLDEsIiIsMSx7InN0eWxlIjp7InRhaWwiOnsibmFtZSI6Imhvb2siLCJzaWRlIjoidG9wIn19fV0sWzMsNCwiIiwxLHsic3R5bGUiOnsidGFpbCI6eyJuYW1lIjoiaG9vayIsInNpZGUiOiJ0b3AifX19XSxbNCw1LCIiLDEseyJzdHlsZSI6eyJ0YWlsIjp7Im5hbWUiOiJob29rIiwic2lkZSI6InRvcCJ9fX1dXQ==
  \begin{tikzcd}
    C & M \\
    B & L \\
    A & K
    \arrow[hook, from=1-1, to=1-2]
    \arrow[hook, from=2-1, to=2-2]
    \arrow[hook, from=3-1, to=3-2]
    \arrow[hook, from=2-1, to=1-1]
    \arrow[hook, from=3-1, to=2-1]
    \arrow[hook, from=3-2, to=2-2]
    \arrow[hook, from=2-2, to=1-2]
  \end{tikzcd}$$
  we have $D_{\wp}(M/L)=D_{\wp}(M/K)\cap\Gal(M/L)$ and $T_{\wp}(M/L)=T_{\wp}(M/K)\cap\Gal(M/L)$ for $p\subseteq A, \pfrak\subseteq B, \wp\subseteq C$ and $\wp|\pfrak, \pfrak|p$. 
\end{proposition}
\section{Lecture 5 -- 25th September 2023}
Let $A$ be a Dedekind domain with fraction field $K$ and $\pfrak\subseteq A$ a prime ideal. Note that $A$ and $A_{\pfrak}$ have the same fraction field $K$ which has all discrete valuations from the primes of $A$. We want to pass to a situation where we completely lose information about primes other than $\pfrak\subseteq A$. We do so via completion. 
\\\\
Let $K$ be a field with discrete valuation $v$ and discrete valuation ring $A$. The ring $A$ has an absolute value by taking $|x|=c^{v(x)}$ for $0<c<1$ setting $|0|=0$. Recall that by \Cref{def:valuation}, we have $|x+y|\leq|x|+|y|$, but the valuation with respect to a prime ideal satisfies a much stronger condition. 
\begin{definition}[Non-Archimedian Valuation]\label{def:nonarchimedian valuation}
  Let $K$ be a field. A non-Archimedian valuation on $K$ is a function $|\cdot|:K\to\RR$ such that:
  \begin{enumerate}[label=(\alph*)]
    \item $|x|\geq0$ for all $x\in K$ and $|x|=0$ if and only if $x=0$, 
    \item $|xy|=|x|\cdot|y|$, 
    \item and $|x+y|\leq\max\{|x|,|y|\}$. 
  \end{enumerate}
\end{definition}
\begin{remark}
  In Serre's text, non-Archimedian valuations are known as ultrametrics. 
\end{remark}
\begin{remark}
  The standard absolute values on number fields, $\RR$, and $\CC$ are Archimedian absolute values. 
\end{remark}
For any field $K$ with an absolute value $|\cdot|$, we have a completion $\widehat{K}$ with respect to the metric $d(x,y)=|x-y|$. These are equivalence classes of Cauchy sequences $\{a_{n}\}_{n\in\NN}$ with $a_{n}\in K$ such that for every $\varepsilon>0$, there exists $N\in \NN$ such that for $n,m\geq N$, $|a_{m}-a_{n}|<\varepsilon$. Two Cauchy sequences are equivalent if they get arbitrarily close together, as in the case of real analysis. We can define $\widehat{K}$ the completion of $K$ with elements Cauchy sequences with the pointwise operations. The additive and multiplicative identities are given by the constant sequences $\{0\}_{n\in\NN}$ and $\{1\}_{n\in\NN}$, respectively. Indeed, this is complete with respect to the absolute value $|\{a_{n}\}_{n\in\NN}|=\lim_{n\to\infty}|a_{n}|$. 
\begin{example}
  Let $K=\QQ$ and $|\cdot|$ the usual (Archimedian) absolute value. We have $\widehat{K}=\RR$. 
\end{example}
\begin{example}
  Let $|\cdot|$ be an absolute value arising from a discrete valuation $v$. We can define $\widehat{v}(x)$ such that for $x\in\widehat{K}$, $|x|=c^{\widehat{v}(x)}$ where $0<c<1$. Note that $c^{\ZZ}$ is discrete so we have $\overline{v}(x)\in\ZZ$ for all $x\in\widehat{K}$ so $\widehat{v}$ is a discrete valuation on $\widehat{K}$ and $\widehat{A}$ is the discrete valuation ring of $\widehat{K}$. In fact, $\widehat{A}$ is the topological closure of $A$ in $\widehat{K}$. 
\end{example}
\begin{definition}[Complete DVR]\label{def: complete DVR}
  Let $\widehat{K}$ be a field complete with respect to the metric induced by its valuation. The discrete valuation ring associated to $\widehat{K}$ is a complete discrete valuation ring. 
\end{definition}
One can also verify that $\widehat{A}$ can be realized as a projective limit 
$$\widehat{A}=\varprojlim_{n}A/(\pi^{n})=\left\{(r_{n})_{n\in\NN}:r_{n}\in A/(\pi^{n}) \text{ s.t. for }m>n, \overline{r_{m}}\equiv r_{n}\pmod{(\pi^{n})}\right\}.$$
Note that in the $p$-adic valuation, things are small if they are divisible by larger powers of $p$. This definition of the completion as inverse limit agrees with that notion. One can verify that $\pi$ is a uniformizer of $\widehat{A}$ and $\widehat{A}_{(\pi)}=\widehat{K}$. 
\begin{proposition}
  Let $S$ be a choice of representatives of $A/(\pi)$ in $A$. Every element of the completion $a\in\widehat{A}$ can be written unqiuely as a convergent power series $a=\sum_{n=0}^{\infty}s_{n}\pi^{n}$ with $s_{n}\in S$
  and every element of the complete field $x\in\widehat{K}$ can be written uniquely as a convergent Laurent series
  $$\sum_{n=m}^{\infty}s_{n}\pi^{n}, m\in\ZZ_{\leq0}.$$
\end{proposition}
\begin{example}
  Let $K=\QQ, A=\ZZ$ and $v$ the $p$-adic valuation with respect to some positive prime $p$. $\widehat{K}=\QQ_{p}$ the $p$-adic numbers with complete discrete valuation ring $\widehat{A}=\ZZ_{p}$, the $p$-adic integers. 
\end{example}
\begin{example}
  Let $K=\FF_{q}(t)$ and $v$ be the $t$-adic valuation, with respect to the prime ideal $(t)$. We have $\widehat{K}=\FF_{q}((t))$ the Laurent series in $t$ and $\widehat{A}=\FF_{q}[[t]]$ the power series in $t$. We can choose $S$ to be elements $\FF_{q}\subseteq\FF_{q}[t]$ so $S$ is an additive and multiplicative monoid and $A/(\pi)=\FF_{q}\subseteq K$. 
\end{example}
Let $A$ be a Dedekind domain with fraction field $K$, $\pfrak\subseteq A$ prime, and $v_{\pfrak}$ the valuation with respect to the prime ideal $\pfrak$. We write $K_{\pfrak}$ be the completion of $K$ with respect to the norm induced by the valuation $v_{\pfrak}$. We have $A_{\pfrak}\subseteq\widehat{A}\subseteq\widehat{K}$ taking $A_{\pfrak}$ to be constant Cauchy sequences. Note that $\widehat{A},\widehat{K}$ are much larger than $A_{\pfrak}$ in the sense that they have solutions to many more equations. These are examples of local fields. \\\\
We return to the setting of extensions. 
\begin{proposition}
  Let $K$ be a field with discrete valuation $v$ and valuation ring $A$ with $K$ complete with respect to the discrete valuation. If further $A$ is a Dedekind domain with fraction field $K$, $L/K$ a finite field extension, and $B$ the integral closure of $A$ in $L$, then $B$ is a discrete valuation ring, a free $A$-module of rank $[L:K]$, and $L$ is complete with respect to the metric from $B$'s valuation. 
\end{proposition}
\begin{proof}
  We show that $B$ is a DVR. The remainder of the proof can be found in \cite[Ch. 2, \S 2, Proposition 3]{Serre}. \\\\
  Let $\wp_{i}$ be the primes of $B$ with valuations $w_{i}$. Note that $\wp_{i}\cap A=(\pi)$ as $(\pi)$ is the only prime ideal of $A$ a DVR. By the equivalence of norms on a finite dimensional vector space over a complete field, each of the $w_{i}$s induces the same topology on $L$. Suppose to the contrary that $B$ is not a DVR, so $w_{1},w_{2}$ are not the same valuation. Pick some $x\in B$ such that $w_{1}(x)>0, w_{2}(x)=0$. In the topology induced by $w_{1}$ on $L$, $\lim_{n\to\infty}x^{n}=0$ but the limit of this sequence, if it exists, would not converge to 0 in the $w_{2}$ topology. This allows us to distinguish between the primes by their associated valuations, a contradiction, so there is only one nonzero prime of $B$ showing that $B$ is a DVR. 
\end{proof}
\begin{remark}
  Note that we required the hypothesis that $K$ complete with respect to the valuation. If $A$ were a Dedekind domain with more than one prime ideal, $A_{\pfrak}$ would be a DVR with the same fraction field as $A$. The integral closure $B$ of $A_{\pfrak}$ in $L$ of $L/K$ a finite extension of fields has primes $\wp_{i}|\pfrak$ and if there is more than one such $\wp_{i}$ then $B$ is not a DVR. 
\end{remark}
\begin{remark}
  In the case of a complete DVR, we need not assume $L/K$ is separable. We use Serre's Hypothesis F where $B$ is finitely generated as an $A$-module where $\sum e_{i}f_{i}=[L:K]$ but in the setup above $i=1$ so $[L:K]=ef$ and there is no splitting of primes. 
\end{remark}
This allows us to deduce the following connection between valuations on $K$ and $L$. 
\begin{corollary}
  For all $x\in L$, $w(x)=\frac{1}{f}v(\Nm_{L/K}(x))$. 
\end{corollary}
\begin{proof}
  This is \cite[Ch. 2, \S 2, Corr. 4]{Serre}. See the discussion following \cite[Ch. 2, \S 2, Proposition 3]{Serre}.
\end{proof}
We return to the setting of extensions of Dedekind domains. Let $A$ be a DVR with fraction field $K$, $L/K$ a finite extension of fields, and $B$ the integral closure of $A$ in $L$. The ring $B$ is a finitely generated $A$-module, but $K$ need note be complete with respect to the metric induced by the discrete valuation. Let $\pfrak$ be a nonzero prime ideal of $A$ and $\wp_{i}\subseteq B$ the primes lying over $\pfrak$. For $x\in K$, we have a valuation $v_{\wp_{i}}$ on $L$ by $e_{i}v_{\pfrak}(x)$, where we say $v_{\wp_{i}}$ prolongs the valuation $v_{\pfrak}$ to $L$ with index $e_{i}$. 
\begin{proposition}
  Let $A$ be a DVR with fraction field $K$, $\pfrak\subseteq A$ prime nonzero, $L/K$ a finite extension of fields, and $B$ the integral closure of $A$ in $L$. If $w$ is a valuation on $L$ prolonging $v_{\pfrak}$, then $w=v_{\wp}$ for some $\wp|\pfrak$. 
\end{proposition}
\begin{proof}
  Let $W$ be the DVR associated to $w$ with maximal ideal $\mfrak$, so $W$ is integrally closed with fraction field $L$. Since $w$ prolongs $v_{\pfrak}$, $A\subseteq W$ and since $W$ is integrally closed, $B\subseteq W$. Let $\wp=\mfrak\cap B$ and note $\wp\cap A=\pfrak$ so $B_{\wp}\subseteq W$ since the denominators outside $\wp$ were outside $\mfrak$. If $W$ contained something with negative $v_{\wp}$ valuation, then $W=L$ a contradiction, so $W=B_{\wp}$ from which it follows that the valuations agree. 
\end{proof}
Let $A$ be a Dedekind domain with fraction field $K$. For $\pfrak\subseteq A$, denote $v_{\pfrak}$ the associated valuation and $\widehat{K}$ the completion of $K$ with respect to the metric induced by $v_{\pfrak}$. Let $L/K$ be a finite extension of fields and $B$ the integral closure of $A$ in $L$. For each $\wp_{i}|\pfrak$, let $\widehat{L_{i}}$ be the completion of $L$ with respect to the metric $v_{\wp_{i}}$ and $\widehat{B_{i}}$ the valuation ring associated to $\widehat{L_{i}}$. Note that from $L$'s point of view, it can't see that all the $\wp_{i}$ lie over the same $\pfrak$. 
\begin{proposition}\label{prop: degree and splitting of completions}
  Let $A$ be a Dedekind domain with fraction field $K$, $\pfrak\subseteq A$ prime nonzero, $L/K$ finite, and $B$ the integral closure of $A$ in $L$. If $\widehat{L_{i}}$ is the completion of $L$ with respect to the metric $v_{\wp_{i}}$ for $\wp_{i}|\pfrak$ and $\widehat{B_{i}}$ the associated valuation rings then 
  \begin{enumerate}[label=(\alph*)]
    \item $[\widehat{L_{i}}:K]=e_{i}f_{\wp_{i}}$, 
    \item $L\otimes_{K}\widehat{K}=L\otimes_{K} K_{\pfrak}=\prod_{i}\widehat{L_{i}}$, 
    \item and $B\otimes_{A}\widehat{A}\cong\prod_{i}\widehat{B_{i}}$. 
  \end{enumerate}
\end{proposition}
\begin{proof}
  This is \cite[Ch. 2, \S 3, Thm. 1]{Serre} for (a) and (b), and \cite[Ch. 2, \S 3, Prop. 4]{Serre} for (c). 
\end{proof}
Of note here is (b), which states that the splitting of primes corresponds to the splitting of the algebra into fields. 
\begin{corollary}
  Assume the conditions of \Cref{prop: degree and splitting of completions}. If further $L/K$ is Galois and $D_{\wp_{i}}$ the decomposition group of $\wp_{i}|\pfrak$, then $\widehat{L_{i}}/\widehat{K}$ is Galois with Galois group $D_{\wp_{i}}$. 
\end{corollary}
\begin{proof}
  The action of $D_{\wp_{i}}$ extends by continuity to $\widehat{L_{i}}$ giving $e_{i}f_{\wp_{i}}=[\widehat{L_{i}}:\widehat{K}]$ automorphisms. 
\end{proof}
\section{Lecture 6 -- 27th September 2023}
We have previously discussed discrete valuation rings. Today we will focus on one important property, Hensel's lemma. 
\begin{lemma}[Hensel I]\label{lem: Hensel I}
  Let $K$ be a complete field with discrete valuation $v$ and $A$ its discrete valuation ring. If $f\in A[x]$ and $a_{0}\in A$ such that $\overline{a_{0}}$ is a simple root of $\overline{f}\pmod{(\pi)}$ then there is a unique root $a\in A$ of $f$ congruent to $\overline{a_{0}}\pmod{(\pi)}$. 
\end{lemma}
The lemma states that if $f$ has a solution modulo $(\pi)$, that solution can be lifted to a solution of $f$ in $A$ itself. This is a very strong statement where we make great use of the characteristics of the non-Archimedian metric. As opposed to the Archimedian setting, Newton's method works fantastically in the $p$-adic topology. 
\begin{proof}[Proof of \Cref{lem: Hensel I}]
  We will construct a Cauchy sequence $\{a_{n}\}_{n\in\NN}$ of elements in $A$ such that $f(a_{n})$ is a root mod $(\pi^{n+1})$ whose limit is $a$. We will then show that $a_{n}$ is unique mod $(\pi^{n+1})$ given the conditions above. 
  \\\\
  Suppose we have $a_{0}\in A$ such that $\overline{a_{0}}$ is a simple root of $f$ modulo $(\pi)$ and set $a_{n+1}$ to be $a_{n}-\frac{f(a_{n})}{f'(a_{n})}$ where $f'(x)$ is the formal derivative of the polynomial $f$. Note that $f'(a_{n})\not\equiv0\pmod{(\pi)}$. If so, we write $\overline{f}=(x-a_{0})\overline{g}$ mod $(\pi)$ and by the product rule we see that $\overline{f}'=\overline{g}+(x-\overline{a_{0}})\overline{g}$ so if $\overline{a_{0}}$ were a root of $\overline{f}'$, it would be a root of $\overline{g}$ implying that it was not a simple root of $f$. 
  \\\\
  Thus $v(f'(a_{0}))=0$, that is $f'(a_{0})\in A^{\times}$ and $a_{n+1}=a_{n}+\frac{f(a_{n})}{f'(a_{n})}\in A$. Recalling the binomial expansion, we write $f(x+y)=f(x)+y f'(x)+y^{2}A[x,y]$ so in our expression we have 
  $$f(a_{n+1})=f\left(a_{n}-\frac{f(a_{n})}{f'(a_{n})}\right)=f(a_{n})-\left(\frac{f(a_{n})}{f'(a_{n})}\right)f'(a_{n})+\dots$$
  where we have that the trailing terms have valuations at least 
  $$v\left(\left(\frac{f(a_{n})}{f'(a_{n})}\right)^{2}\right)\geq 2\cdot v(f(a_{n}))\geq 2(n+1)\geq n+2$$
  so $f(a_{n+1})\equiv 0\pmod{(\pi^{n+2})}$ demonstrating existence of the Cauchy sequence. 
  \\\\
  To demonstrate uniqueness, assume that $a_{n+1}$ is unqiue modulo $(\pi^{n+1})$ so if $f(a_{n}+h\pi^{n+1})=0$ then using our expansion above we have 
  $$f(a_{n})+h\pi^{n+1}f'(a_{n})+\dots$$
  where the trailing terms have valuation at least $n+2$ which determines $h$ modulo $(\pi)$ implying the uniqueness of $a_{n+1}$ modulo $(\pi^{n+2})$. This implies that the sequence is Cauchy, and taking $a=\lim_{n\to\infty}a_{n}$, we have $f(a)=0$ and $a$ is unique since each $a_{n}$ is determined uniquely modulo $(\pi^{n})$ for all $n$. 
\end{proof}
Note that Hensel's lemma comes in many forms and can be made more general in at least two ways which we now state. 
\begin{lemma}[Hensel, II]\label{lem: Hensel II}
  Let $f\in A[x]$ and $a_{0}\in A$ such that $\overline{a_{0}}$ is a simple root of $\overline{f}\pmod{(\pi)}$. If $v(f(a_{0}))>2\cdot v(f'(a_{0}))$ then there is a root $a\in A$ of $f(x)$. 
\end{lemma}
\begin{lemma}[Hensel, III]\label{lem: Hensel III}
  Let $f\in A[x]$ be a monic polynomial. If $\overline{f}$ factors as $g_{0}h_{0}$ for $g_{0},h_{0}\in (A/(\pi))[x]$ with $g_{0},h_{0}$ monic and relatively prime, then $f=gh$ for $g,h\in A[x]$ such that $\overline{g}=g_{0}, \overline{h}=h_{0}$. 
\end{lemma}
There is also a version that implies \Cref{lem: Hensel I,lem: Hensel II,lem: Hensel III} all together. Furthermore, $f$ need not be monic as stated in the version of \cite[Ch. 2, Thm. 4.6]{Neukirch}. Hensel's lemma is a very extensive statement stating that the vast majority of the time the factorization of polynomials in a complete DVR is determined by its factorization in the residue field. 
\begin{example}
  Consider $x^{p-1}-1\in \ZZ_{p}[x]$ a polynomial over the $p$-adic integers. THis factors into distinct linear factors $(\ZZ_{p}/p\ZZ_{p})\cong(\ZZ/p\ZZ)\cong\FF_{p}$. Applying \Cref{lem: Hensel II} repeatedly $p-2$ times, this implies that $x^{p-1}-1$ factors into distinct linear factors in $\ZZ_{p}$. That is, $\ZZ_{p}$ contains $p-1$ $(p-1)$th roots of unity. 
\end{example}
This gives a multiplicatively closed set of representatives for the residue classes for the $(p-1)$th roots of unity. Recall here that $\ZZ_{p}$ (uncountable) is much larger than the localizations $\ZZ_{(p)}$ (countable). Similarly if $A/(\pi)\cong\FF_{q}$ for $q$ some prime power then $K=K(A)$ has all $(q-1)$th roots of unity. 
\\\\
We are now able to define local fields. 
\begin{definition}[Local Field]\label{def: local field}
  Let $K$ be a complete field with discrete valuation $v$ and $A$ its discrete valuation ring. $K$ is a local field if and only if $|A/(\pi)|<\infty$, that is, it has finite residue field. 
\end{definition}
Local fields arise in two families: the mixed characteristic case and the equicharacteristic case. 
\begin{definition}[Mixed Characteristic Local Field]\label{def: mixed characteristic local field}
  Let $K$ be a local field and $A$ its discrete valuation ring with uniformizer $\pi$. $K$ is a mixed characteristic local field if $\mathrm{char}(K)=0$. 
\end{definition}
\begin{remark}
  In the setting of \Cref{def: mixed characteristic local field}, evidently $\mathrm{char}(A/(\pi))=p$ for some prime $p$ since $K$ is a local field. 
\end{remark}
In the setting of \Cref{def: mixed characteristic local field}, $K$ is characteristic 0 so its discrete valuation ring $A$ contains the integers $\ZZ$. So for $\mathrm{char}(A/(\pi))=p$, $p\in A$ where $v(p)\geq 1$ say $v(p)=e$. We thus write $p=\pi^{e}u$ for some unit $u\in A^{\times}$. This leads to the definition of the absolute ramification index. 
\begin{definition}[Absolute Ramification Index]\label{def: absolute ramification index}
  Let $K$ be a local field of characteristic 0 with discrete valuation ring $A$ and $\mathrm{char}(A/(\pi))=p$. The absolute ramification index is $v(p)$. 
\end{definition}
Note that in the setting above $p$ has positive valuation implying that $A$ contains the $p$-adic integers as well. Let $f=[(A/(\pi)):\ZZ/p\ZZ]$. One can show that $A$ is a free $\ZZ_{p}$-module of rank $ef$ and $[K:\QQ_{p}]=ef$ as well. This is an extension of Dedekind domains where $e$ is the exponent of $(\pi)$ lying over $(p)\subseteq\ZZ$.
\begin{example}
  Let $K/\QQ$ be a finite extension of fields unramified at $(p)$ and $\pfrak$ a prime of $\Ocal_{K}$ over $(p)$. The completion of $K_{\pfrak}$ with respect to the $\pfrak$-adic valuation $v_{\pfrak}$ is an absolutely unramified local field.
\end{example} 
This seems to suggest that absolutely unramified local fields are plentiful, but the following result shows otherwise. 
\begin{theorem}\label{thm: unique mixed characteristic extension}
  For every finite field $\FF_{q}$ of characteristic $p$, there is a complete local field $K$ of characteristic 0 which is absolutely unramified with residue field $\FF_{q}$. 
\end{theorem}
\begin{proof}[Proof of Uniqueness]
  Let $q=p^{f}$ and $\overline{\theta}$ generate $\FF_{q}$ over $\ZZ/p\ZZ$ whose existence is given by the primitive element theorem over finite fields with minimal polynomial $\overline{h}\in \FF_{p}[x]$ of degree $f$. Let $A,A'$ be complete DVRs of characteristic 0 absolutely unramified of characteristic 0 with residue fields $\FF_{q}$. Denoting $K,K'$ the fraction fields of $A,A'$, respectively, $[K:\QQ_{p}]=[K':\QQ_{p}]=f$. Since $\overline{h}$ is minimal, $\overline{\theta}$ is a simple root and applying \Cref{lem: Hensel I}, we can lift it to $\theta\in A$ whose minimal polynomial is $g\in\ZZ_{p}[x]$. Note that $\overline{g}(\overline{\theta})=0$ modulo $(\pi)$ so $\overline{h}|\overline{g}$ so the degree of $g$ is at least $f$, but $K,K'$ are degree $f$ extensions so the degree of $g$ is $f$ as well and $\overline{h}=\overline{g}$ and $K\cong \QQ_{p}[x]/(g(x))$. \\\\
  In $K'$ consider $g(x)\in A'[x]$ and note that $g$ splits into distinct linear factors over $\FF_{q}[x]$ since $g$ has root $\overline{\theta}\in\FF_{q}$ modulo $(\pi')$ where $\pi'$ is the uniformizer of the DVR $A'$ and $\FF_{q}/\FF_{p}$ is Galois. Applying \Cref{lem: Hensel II}, $g$ factors into distinct linear factors in $A'[x]$ so $K'$ contains a root of $g$ and $[K':\QQ_{p}]=\deg g$. But $g$ was irreducible over $\ZZ_{p}$ as $\overline{g}$ was irreducible over $\FF_{p}[x]$ so $K'\cong\QQ_{p}[x]/(g(x))$ giving the isomorphism.
\end{proof}
The proof of existence is \cite[Ch. II, \S 5, Thm. 3]{Serre}. 
\begin{example}
  Let $K/\QQ$ be a finite extension of fields with $\pfrak_{1},\pfrak_{2}$ over $(p)\subseteq\ZZ$. If $e_{1}=f_{\pfrak_{1}}=e_{2}=f_{\pfrak_{2}}=1$ then 
  $$[K_{\pfrak_{1}}:\QQ_{p}]=e_{1}f_{\pfrak_{1}}=1=e_{2}f_{\pfrak_{2}}=[K_{\pfrak_{2}}:\QQ_{p}]$$
  and $K_{\pfrak_{1}}\cong K_{\pfrak_{2}}\cong\QQ_{p}$. If $e_{1}=e_{2}=1$ and $f_{\pfrak_{1}}=f_{\pfrak_{2}}$ then $K_{\pfrak_{1}}\cong K_{\pfrak_{2}}$
\end{example}
\section{Lecture 7 -- 2nd October 2023}
Recall \Cref{thm: unique mixed characteristic extension} from the previous lecture. We begin with a generalization of this theorem. 
\begin{theorem}
  Let $K$ be a complete local field with discrete valuation ring $A$ and $A/(\pi)\cong\FF_{q}$. There is a unique unramified extension $L/K$ of degree $[L:K]=f$ with residue field $\FF_{q^{f}}$. 
\end{theorem}
Last time, we saw what the characteristic 0 local fields could be in terms of $p$-adic fields using Hensel's Lemma (\Cref{lem: Hensel I,lem: Hensel II,lem: Hensel III}). However, Hensel's lemma works over any complete local field so the same proof works in this more general setting. We will soon try to understand extensions of arbitrary local fields. Recall that in the setting of a complete local field $K$ with discrete valuation ring $A$ unramified extensions of $K$ correspond bijectively to extensions of the residue field $A/(\pi)$. In the complete setting, the single decomposition group is $\Gal(L/K)$ and we have an isomorphism $\Gal(L/K)\cong\Gal((B/(\pi_{B}))/(A/(\pi_{A})))$. 
\\\\
Let $K$ be a characteristic 0 local field. From \Cref{thm: unique mixed characteristic extension}, $K$ is a finite extension of $\QQ_{p}$. If further $e>1$, we can consider the maximal unramified extension $K_{T}$ for $T$ the inertia group. So we have a tower of fields 
$$% https://q.uiver.app/#q=WzAsMyxbMCwwLCJLIl0sWzAsMSwiS197VH0iXSxbMCwyLCJcXFFRX3twfSJdLFsxLDIsIlxcbWF0aHJte3VucmFtZmllZH0iLDEseyJzdHlsZSI6eyJoZWFkIjp7Im5hbWUiOiJub25lIn19fV0sWzAsMSwiXFx0ZXh0e3JhbWlmaWVkfSIsMSx7InN0eWxlIjp7ImhlYWQiOnsibmFtZSI6Im5vbmUifX19XV0=
\begin{tikzcd}
	K \\
	{K_{T}} \\
	{\QQ_{p}}
	\arrow["{\mathrm{unramfied}}"{description}, no head, from=2-1, to=3-1]
	\arrow["{\text{ramified}}"{description}, no head, from=1-1, to=2-1]
\end{tikzcd}$$
where $K/K_{T}$ is an extension of degree $[K:K_{T}]=e$. By Krasner's Lemma \cite[\href{https://stacks.math.columbia.edu/tag/0BU9}{0BU9}]{stacks-project}, there are only finitely many extensions for given $e$. This gives us great control over the types of local fields that arise. We will soon see the equal characteristic case for local fields, which will eventually lead to a complete classification. 
\\\\
Let us now consider the equal characteristic case. 
\begin{definition}[Equal Characteristic Local Fields]\label{def: equal characteristic local field}
  Let $K$ be a local field and $A$ its discrete valuation ring. $K$ is an equal characteristic local field if $\mathrm{char}(A/(\pi))=\mathrm{char}(K)$. 
\end{definition}
As in the case of \Cref{thm: unique mixed characteristic extension}, we have a classification theorem in the equicharacteristic case. 
\begin{theorem}
  If $K$ is a local field of characteristic $p$ with residue field $A/(\pi)\cong\FF_{q}$ for $q$ a $p$-th power, $A\cong\FF_{q}[[T]]$ the power series in $\FF_{q}$ and $K\cong\FF_{q}((T))$ the Laurent series in $\FF_{q}$. 
\end{theorem}
\begin{proof}
  We will find a set $S\subseteq A$ of representatives of the residue field that is additively and multiplicatively closed. The theorem will then follow from 
\end{proof}
\newpage
\printbibliography
\end{document}